Blame and Coercion: Together Again for the Second Time

### 2019-10-03 paper.tex
head of section 2

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{stmaryrd}
\usepackage{todonotes}

\newtheorem{theorem}{Theorem}[section]
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\stxrule}[3]{#1 & ::= & #3 & \text{#2}\\}
\newcommand{\funrule}[3]{#1 & = & #2 & #3\\}
\newcommand{\comprule}[4]{#1 & \fatsemi^l & #2 & = & #3 & #4 \\}
\newcommand{\plus}[0]{+}

@@

\citet{siek2010threesomes} proposes another approach to combine blame 
tracking and space efficiency. Their solution is based on Threesome Calculus, 
an novel alternative to Coercion Calculus. The key idea is to represent casts 
with threesomes. Threesomes, like coercions, can be composed. There is no 
separate normalization for threesome because every threesome is normalized. 
They prove that their Threesome Calculus bisimulate 
\citeauthor{siek2009exploring}'s Coercion Calculus and
\citeauthor{wadler2009well}'s Blame Calculus.

\citet{siek2012interpretations} introduces a $ lazy D $  Blame Calculus. The 
relation between this new Blame Calculus with the $ lazy D $ 
Coercion Calculi, however, is not proved.

Following \citet{siek2010threesomes}, \citet{garcia2013calculating} 
implement all other blame strategies with Threesome Calculus. 
He claims that 
coercion with labels is easy to understand but hard to implement, 
and that 
threesome with labels, however, is easy to implement but hard to understand.
His claim is later affirmed by the group of people who develop threesome 
\citet{siek2015blame}. 
The connection 
between his Threesome Calculi and \citeauthor{siek2009exploring}'s Coercion 
Calculi are established by the fact that the former are derived from the latter.
The connection between these calculi and Blame Calculus, however, 
is still unclear.

@@

Last year, \citet{kuhlenschmidt2018efficient} present Grift, a space-efficient 
and blame-tracking compiler for a gradually typed language of the same name. 
This implementation is based on the $ Lazy D $ Coercion Calculus.
Their result suggests that implementing coercion is practical.

@@

\begin{itemize}
	\item Is the $ Lazy D $ coercion an ideal cast representation?
	\item Does the $ Lazy D $ Calculi bi-simulate each other?
\end{itemize}

@@

syntax, consistency, shallow consistency ...

@@

\begin{theorem}[Hyper-coercion bi-simulates cast]	
	$ e \Downarrow_B o $ if and only if $ e \Downarrow_H o $
\end{theorem}

>>>>

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{stmaryrd}
\usepackage{todonotes}
\usepackage{amsthm}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{semantic}
\usepackage{hyperref}

\newtheorem{theorem}{Theorem}[section]
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\stxrule}[3]{#1 & ::= & #3 & \text{#2}\\}
\newcommand{\stxrulecont}[1]{& | & #1 & \\}
\newcommand{\funrule}[3]{#1 & = & #2 & #3\\}
\newcommand{\comprule}[4]{#1 & \fatsemi^l & #2 & = & #3 & #4 \\}
\newcommand{\plus}[0]{+}
\newcommand{\judgetype}[3]{#1 \vdash #2 : #3}

@@

\citet{siek2010threesomes} proposes another approach to combine blame 
tracking and space efficiency. Their solution is based on Threesome Calculus, 
an novel alternative to Coercion Calculus. The key idea is to represent casts 
with threesomes. Threesomes, like coercions, can be composed. There is no 
separate normalization for threesome because every threesome is normalized. 
They prove that their Threesome Calculus bisimulate 
\citet{siek2009exploring}'s Coercion Calculus and
\citet{wadler2009well}'s Blame Calculus.

\citet{siek2012interpretations} introduces a $ lazy D $  Blame Calculus. They 
conjecture that this calculus bi-simulate the $ lazy D $ Coercion Calculi.

Following \citet{siek2010threesomes}, \citet{garcia2013calculating} 
implement all other blame strategies with Threesome Calculus. 
He claims that 
coercion with labels is easy to understand but hard to implement, 
and that 
threesome with labels, however, is easy to implement but hard to understand.
His claim is later affirmed by the group of people who develop threesome 
\citet{siek2015blame}. 
The connection 
between his Threesome Calculi and \citet{siek2009exploring}'s Coercion 
Calculi are established by the fact that the former are derived from the latter.
The connection between these calculi and Blame Calculus, however, 
is still unclear.

@@

Last year, \citet{kuhlenschmidt2018efficient} present Grift, a space-efficient 
and blame-tracking compiler for a gradually typed language of the same name. 
This implementation is based on the $ Lazy D $ Coercion Calculus.
Their result suggests that implementing coercion is practical. And this is the 
first time product types is considered. Their treatment to products in 
coercions, however, is shown incorrect. **

\todo[inline]{kc: ** I want to cite the GitHub issue page here.}

@@

\begin{itemize}
	\item Is the $ Lazy D $ coercion an ideal cast representation?
	\item Does the $ Lazy D $ Coercion Calculus bi-simulate the Blame Calculus?
\end{itemize}

@@

\begin{figure}
	\[
	\begin{array}{lclr}
	\stxrule{S,T}{types}{
		\star \mid{}
		P
	}
	\stxrule{P,Q}{pre-types}{
		\iota \mid{}
		T_1 \rightarrow T_2 \mid{}
		T_1 \times T_2 \mid{}
		T_1 \plus T_2 \mid{}
	}
%	\stxrule{x,y,z}{variables}{\dots}
%	\stxrule{l}{labels}{\dots}
%	\stxrule{k}{constants}{sole}
	\stxrule{e}{terms}{
		x \mid{}
		k \mid{}
		\lambda^{S\rightarrow{}T}x.t \mid{}
		e_1 \; e_2 \mid{}
		\langle T \Leftarrow^l S \rangle t \mid{}
		\mathtt{blame} \; l
	}
	\stxrulecont{
		\mathtt{cons}\; e_1 \; e_2 \mid{}
		\mathtt{car}\; e \mid{}
		\mathtt{cdr}\; t
	}
	\stxrulecont{
		\mathtt{inl} \; e \mid{}
		\mathtt{inr} \; e \mid{}
		\mathtt{case}\; e_1 \; e_2 \; e_3
	}
	\stxrule{o}{observations}{
		k \mid{}
		\mathtt{fun}\mid{}
		\mathtt{cons}\mid{}
		\mathtt{inl}\mid{}
		\mathtt{inr}\mid{}
		\mathtt{blame}\; l
	}
	\end{array}
	\]

	\fbox{$ S \sim T $}
	\begin{gather*}
	\inference{}{
		\star \sim \star
	} \quad
	\inference{}{
		\star \sim P
	} \quad
	\inference{}{
		P \sim \star
	} \\
	\inference{}{
		\iota \sim \iota
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \rightarrow T_1 \sim S_2 \rightarrow T_2
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \times T_1 \sim S_2 \times T_2
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \plus T_1 \sim S_2 \plus T_2
	}
	\end{gather*}

	\fbox{$ S \smile T $}
	\begin{gather*}
	\inference{}{
		\star \smile \star
	} \quad
	\inference{}{
		\star \smile P
	} \quad
	\inference{}{
		P \smile \star
	} \\
	\inference{}{
		\iota \smile \iota
	} \quad
	\inference{}{
		S_1 \rightarrow T_1 \smile S_2 \rightarrow T_2
	}
	\inference{}{
	S_1 \times T_1 \smile S_2 \times T_2
	}
	\inference{}{
	S_1 \plus T_1 \smile S_2 \plus T_2
	}
	\end{gather*}

	\fbox{$ \Gamma \vdash e \vdash T $}
	\begin{gather*}
		\inference{
			S \sim T & \Gamma \vdash e : S 
		}{
			\judgetype{\Gamma}{\langle T \Leftarrow^l S \rangle t}{T}
		} \quad
		\inference{
		}{
			\judgetype{\Gamma}{\mathtt{blame} l}{T}
		}
	\end{gather*}

	\caption{Blame Calculus and its static semantics}
	\label{fig:blame-static}
\end{figure}

\figref{fig:blame-static} defines the form of blame calculus and its static 
semantics. It is little changed from previous definitions. 
The dynamic semantics of Blame Calculi depend on blame strategies, so we defer 
them to sub-sections.

Blame Calculus is based on Simply Typed Lambda Calculus with sum types and 
product types ($ \mathtt{STLC+} $). 
Let $ S,T $ range over types. A type is either the dynamic type $ \star $
(a.k.a. $ \mathtt{Dyn} $, $ \mathbb{?} $, or $ \mathtt{Unknown} $), 
or a pre-type. 
Let $ P,Q $ range over pre-types. Every pre-type is a type with a 
traditional type constructor at the top.

$ S \sim T $ reads $ S $ and $ T $ are consistent. Two types are consistent if 
one of them is $ \star $, or they have the same top-most type constructor and 
the corresponding sub-parts are consistent. Consistency is reflexive and 
symmetric, but not transitive.

$ S \smile T $ reads $ S $ and $ T $ are shallowly-consistent. Two types are 
shallowly-consistent if one of them is $ \star $, or they have the same 
top-most type constructor. 
Shallow-consistency is also reflexive, symmetric, and not transitive.

Let $ e $ ranges over terms. Unlike $ \mathtt{STLC+} $, we annotate the 
co-domain of lambda abstractions explicitly and add casts and blames.

@@

\begin{theorem}[Hyper-coercion bi-simulates cast]	
	$ t \Downarrow_B o $ if and only if $ t \Downarrow_H o $
\end{theorem}
